---
title: "Document Title Trial"
subtitle: "This template supports subtitles also."
author: "Alejandro Abraham"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_document:
        toc: true
        number_sections: false
        css: style_html.css
        template: template_banner.html
---


```{r setup, include=FALSE}

# Global options
knitr::opts_chunk$set(echo=FALSE,
	                  cache=FALSE,
                      prompt=FALSE,
                      comment=NA,
                      message=FALSE,
                      warning=FALSE,
                      fig.align = 'centre',
                      fig.width = 10,
                      fig.retina = 3)

# Kable Options
options(knitr.kable.NA = '')

# Load the innteamUtils and framework packages
library(innteamUtils)
library(data.table)
library(reactable)
library(ggplot2)
library(plotly)

### ggplot2 global theme
theme_set(theme_innteam())

### reactable global theme
options(reactable.theme = react_format)

```


Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

### Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

### ggplot2 Example

```{r, echo=TRUE}

dt <- as.data.table(mtcars)

dt[, cyl := as.factor(cyl)] %>% 
    ggplot(
        aes(
           x = hp,
           y = mpg,
           color = cyl)) + 
    geom_point() +
        theme_innteam() + 
            scale_color_brewer(type = 'qual', palette = 6) +
                 labs(title = 'Another mtcars Plot',
                      subtitle = 'But this time it is a data.table',
                      caption = 'The theme is coherent with the overall theme in the package')

```

### Plotly Example

```{r, echo=TRUE}

dt %>%
  plot_ly(x = ~hp,
          y = ~mpg,
          color = ~cyl,
          type = "scatter") %>% 
    plotly_innteam(plot_title = 'This is my title',
                     plot_subtitle = 'Here we will see a subtitle',
                     plot_caption = 'Source: innteamUtils package',
                   plot_xaxis = 'LifeExp',
                   plot_yaxis = 'Population',
                     width = 900, height = 500)

```


You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `reactable()`.

```{r, echo=TRUE, results='asis'}
reactable(iris, highlight = TRUE, filterable = TRUE, resizable = TRUE, wrap = FALSE)
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))


```{r, echo=TRUE}
x <- seq(1:5)
x
```

