---
title: "Datasets for Common Activities"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Datasets for Common Activities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  comment = "#",
    error = FALSE,
     tidy = FALSE,
    cache = FALSE,
 collapse = TRUE
)
```

```{r setup}
library(innteamUtils)
```



## Introduction

The package contains several datasets that are useful in daily activities and have been processed from different sources or for practical usefulness. 

## Table of Contents

Currently the following sets are available:  

- `geo_italia`: a long-format table containing **4 NUTS levels** from EUROSTAT for Italy.  
- `decodifica_comuni`: a wide-format table containing *old* and *new* Italian municipalities (and higher level denomination) names.  
- `codice_ateco`: a long-format table containing **all NACE Rev.2 levels** in Italian language.  
- `agenzie_provincia`: simple table containing the number of insurance agencies and subagents by Italian province. Strictly for internal use only.



## Preview

### Italian Geographic Levels

The dataset has `r nrow(geo_italia)` observations and `r ncol(geo_italia)` columns.  
Curated by Fabio Sgobio, Alejandro Abraham. 

```{r}
head(geo_italia)
```


### Decodifica Comuni

The dataset has `r nrow(decodifica_comuni)` observations and `r ncol(decodifica_comuni)` columns.   Curated by Fabio Sgobio.

```{r}
head(decodifica_comuni)
```


### Codice Ateco 2021

The dataset has `r nrow(codice_ateco)` observations and `r ncol(codice_ateco)` columns.  
Curated by Fabio Sgobio, Alejandro Abraham. 

```{r}
head(codice_ateco)
```



### Agenzie e Subagenti per Provincia Italiana

The dataset has `r nrow(agenzie_provincia)` observations and `r ncol(agenzie_provincia)` columns.  
Curated by Fabio Sgobio, Carlo Cardinali, Alejandro Abraham

```{r}
head(agenzie_provincia)
```


## Usage Example  

Even if you load the `innteamUtils` package into your R session, datasets are available but not loaded into your current enviornment. To load them you need to call them explicitly with the `data()` function.  For example:  

```{r echo=TRUE, eval=TRUE}
# Call the dataset: don't worry, it is a "promise" until you use it.
data("agenzie_provincia")

# Calculate the Mean 
agenzie_provincia[, lapply(.SD, mean), .SDcols = c('agenzie', 'subagenti')]


```

