---
title: "Innovation Team Standard R Theming kit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Innovation Team Standard R Theming kit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  comment = "#",
    error = FALSE,
     tidy = FALSE,
    cache = FALSE,
 collapse = TRUE
)
```

```{r setup}
library(innteamUtils)
```

## Introduction

A great deal of the current package is dedicated to theming and reporting means. Communication is a central skill in nowadays data analysis, therefore the following achievements are seeked:  
- Standardization: sign of professionalism, makes a greater visual impact and allows better storytelling.  
- Time Saving: standard themes provide the code and arrangements, saving typing and thinking on visual aspects.  
- Spillover Effect: more time and energy available for analytical aspects.  

## Theme Elements:  

For these purposes, the package contains:  
- **Report PDF/HTML ready**: A curated skeleton easy and lightweight to rendering thanks to `pagedown`, with front and back cover. Available in the **Rmarkdown Templates**.  
- **Quick HTML Document**: a simple HTML document, lightweight and concise. Available in the **Rmarkdown Templates**.   
- `ggplot2` **theme**: a ready out-of-the-box theme for plots via theme function.  
    - `+ theme_innteam()`  
- **PNG export**: optional with the `ggplot2` theme, this option allows to print plots as PNG's with a brand bottom banner.  
    - `export_png()`

